Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-03-12 Grassmannian_cells.tex
Small section 5.2 edit
cleaned up references, added acknowledgements

\author{Dylan Rupel}
\address[Dylan Rupel]{University of Notre Dame, Department of Mathematics, Notre Dame, IN 46556, USA}
\email{drupel@nd.edu}
\author{Thorsten Weist}
\address[Thorsten Weist]{}
\email{weist@uni-wuppertal.de}

@@

The results of \cite{rupel} giving a combinatorial construction of counting polynomials for quiver Grassmannians of preprojective/preinjective representations of generalized Kronecker quivers $K(n)$ suggest the following.
\begin{conjecture}
  There exists a torus action on $\Gr_\bfe(P_{m+1}^V)$ such that the affine cell corresponding to a compatible pair $(S_H,S_V)$ has dimension...
\end{conjecture}
\sayD{I need to include a conjecture about the dimensions of cells coming from the combinatorial construction of counting/Poincare polynomials}
\sayT{Insert something about Dyck path combinatorics and the correspondence here}
\sayT{maybe we can erase some parts which show up twice in the introduction?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quiver Covering Theory}
\label{sec:covering}
\noindent
We refer to \cite{gab} for an introduction to covering theory.
Let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:s(\alpha)\to t(\alpha)$.
A $\CC$-representation $X$ of $Q$ consists of a collection of $\CC$-vector spaces $X_i$ for $i\in Q_0$ and a collection of $\CC$-linear maps $X_\alpha:X_{s(\alpha)}\to X_{t(\alpha)}$ for $\alpha\in Q_1$.
Given $\CC$-representations $X$ and $Y$ of $Q$, a morphism $f:X\to Y$ is a collection of $\CC$-linear maps $f_i:X_i\to Y_i$ for $i\in Q_0$ satisfying $f_{t(\alpha)}\circ X_\alpha=Y_\alpha\circ f_{s(\alpha)}$ for each $\alpha\in Q_1$.
We write $\rep Q$ for the hereditary abelian category of finite-dimensional $\C$-representations of $Q$ and we assume in the following that all representations are finite-dimensional.

@@

For $m\ge1$ and $1\le i\le n-1$, we write $D_{m,i}$ for the $i$-th copy of $D_m$ inside $D_{m+1}$.
Note that for $1\le r\le n-1$, the maximal Dyck paths $D_{m,i}$, $r+1\le i\le n$, naturally identify with subpaths of $D_{m+1}^{[r]}$.
Extending the notation above, for $m\ge2$ and $1\le r\le n-1$, we write $D_{m,i}^{[r]}$ for the Dyck path obtained by removing the first $r$ copies of $D_{m-1}$ from $D_{m,i}$.
\begin{remark}
  For notational convenience, we also set $D_{m,n}^{[1]}:=D_{m+1}^{[n-1]}$ even though there is no maximal Dyck path $D_{m,n}$ identifying with a copy of $D_m$ inside $D_{m+1}$, such notation is justified by Theorem~\ref{th:dyck path recursion}.
  This allows to write $D_{m,n}^{[r]}$ for $1\le r\le n-1$ for the terminal subpaths of $D_{m+1}$.
  This should be compared with Corollary~\ref{cor:truncated preprojective isomorphism} and Lemma~\ref{le:truncated quotients}.
\end{remark}

@@

\sayT{references to be unified}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{10}

\bibitem{ass} 
  I.~Assem, D.~Simson, A.~Skowronski: Elements of the Representation Theory of Associative Algebras. Cambridge University Press, Cambridge 2007.
\bibitem{ars} 
  M.~Auslander, I.~Reiten, S.~O.~Smalo: Representation theory of Artin algebras {\bf 36}. Cambridge University Press, Cambridge 1997.
\bibitem{bb} 
  A.~Bia\l{}ynicki-Birula: Some theorems on actions of algebraic groups. Annals of Mathematics \textbf{98}, 480-497 (1973).
\bibitem{bgp} 
  I.~N.~Bernstein, I.~M.~Gelfand, V.~A.~Ponomarev: Coxeter functors, and Gabriel's theorem. Russian Mathematical Surveys \textbf{28}(2), 17-32 (1973).
\bibitem{brenner-butler} 
  S.~Benner, M.~C.~R.~Butler, The equivalence of certain functors occuring in the representation theory of artin algebras and species, J. London Math. Soc., 14 (1976), 183-187.
\bibitem{cc} 
  P.~Caldero, F.~Chapoton: Cluster algebras as {H}all algebras of quiver representations. Commentarii Mathematici Helvetici \textbf{81}(3), 595-616 (2006).
\bibitem{ck} 
  P.~Caldero, B.~Keller: From triangulated categories to cluster algebras II.  Ann. Sci. \'Ecole Norm. Sup. (4) \textbf{39} (2006), no. 6, pp.~983--1009.
\bibitem{ce} 
  G.~Cerulli Irelli, F.~Esposito: Geometry of quiver Grassmannians of Kronecker type and applications to cluster algebras. Algebra \&  Number Theory, 5(6), 777-801.
\bibitem{cefr} 
  G.~Cerulli Irelli, F.~Esposito, H.~Franzen, M.~Reineke: Topology of Quiver Grassmannians. Preprint 2017.
\bibitem{cr} 
  P.~Caldero, M.~Reineke: On the quiver Grassmannian in the acyclic case. Journal of Pure and Applied Algebra \textbf{212}(11), 2369-2380 (2008).
%\bibitem{cj} Crawley-Boevey, W., Jensen, Bernt Tore: A note on sub-bundles of vector bundles. Glasgow Mathematical Journal \textbf{48}, 459-462 (2006).
\bibitem{gab} 
  P.~Gabriel: The universal cover of a finite-dimensional algebra. Representations of algebras. Lecture Notes in Mathematics {\bf 903}, 68-105 (1981).
\bibitem{hr} 
  D.~Happel, C.~M.~Ringel: Tilted Algebras. Transactions of the American Mathematical Society {\bf 274}, no.2, 399-443 (1982).
\bibitem{llz} 
  K.~Lee, L.~Li, A.~Zelevinsky: Greedy elements in rank 2 cluster algebras. Selecta Math. \textbf{20} (2014) pp.~57--82.
\bibitem{lw} 
  O.~Lorscheid, T.~Weist: Representation type via Euler characteristics and singularities of quiver Grassmannians. Preprint 2017, arXiv:1706.00860.
\bibitem{rin1} 
  C.~M.~Ringel. Exceptional modules are tree modules. \textit{Linear algebra and its Applications}, 275/276:471-493, 1998.
\bibitem{rin} 
  C.~M.~Ringel: Reflection functors for hereditary algebras. Journal of the London Mathematical Society (2) {\bf 21}, no. 3, 465-479 (1980).
%\bibitem{mats} Matsumura, H.: Commutative ring theory. 
\bibitem{rupel} 
  D.~Rupel: Rank Two Non-Commutative Laurent Phenomenon and Pseudo-Positivity, arXiv.
\bibitem{sch} 
  A.~Schofield: General representations of quivers. Proceedings of the London Mathematical Society (3) \textbf{65}(1), 46-64 (1992).
\bibitem{wei} 
  T.~Weist: Localization of quiver moduli spaces. Representation Theory \textbf{17}(13), 382-425 (2013).
\bibitem{wei2} 
  T.~Weist: Tree modules. Bulletin of the London Mathematical Society \textbf{44}(5), 882-898 (2012).
%\bibitem{wei3} Weist, T.: On the recursive construction of indecomposable quiver representations. Journal of Algebra \textbf{443}, 49-74 (2015).
\end{thebibliography}

>>>>

\author{Dylan Rupel}
\address[Dylan Rupel]{University of Notre Dame, Department of Mathematics, Notre Dame, IN 46556, USA}
\email{drupel@nd.edu}
\author{Thorsten Weist}
\address[Thorsten Weist]{Bergische Universit\"at Wuppertal, Gau\ss str.\ 20, 42097 Wuppertal, Germany}
\email{weist@uni-wuppertal.de}

@@

The results of \cite{rupel} giving a combinatorial construction of counting polynomials for quiver Grassmannians of preprojective/preinjective representations of generalized Kronecker quivers $K(n)$ suggest the following.
\begin{conjecture}
  There exists a torus action on $\Gr_\bfe(P_{m+1}^V)$ such that the affine cell corresponding to a compatible pair $(S_H,S_V)$ has dimension...
\end{conjecture}
\sayD{I need to include a conjecture about the dimensions of cells coming from the combinatorial construction of counting/Poincare polynomials}
\sayT{Insert something about Dyck path combinatorics and the correspondence here}
\sayT{maybe we can erase some parts which show up twice in the introduction?}
\subsection*{Acknowledgements}We would like to thank Giovanni Cerulli Irelli, Hans Franzen, Oliver Lorscheid and Markus Reineke for very fruitful discussions related to this project.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quiver Covering Theory}
\label{sec:covering}
\noindent
We refer to \cite{gab} for an introduction to covering theory.
Let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:s(\alpha)\to t(\alpha)$.
A $\CC$-representation $X$ of $Q$ consists of a collection of $\CC$-vector spaces $X_i$ for $i\in Q_0$ and a collection of $\CC$-linear maps $X_\alpha:X_{s(\alpha)}\to X_{t(\alpha)}$ for $\alpha\in Q_1$.
Given $\CC$-representations $X$ and $Y$ of $Q$, a morphism $f:X\to Y$ is a collection of $\CC$-linear maps $f_i:X_i\to Y_i$ for $i\in Q_0$ satisfying $f_{t(\alpha)}\circ X_\alpha=Y_\alpha\circ f_{s(\alpha)}$ for each $\alpha\in Q_1$.
We write $\rep Q$ for the hereditary abelian category of finite-dimensional $\C$-representations of $Q$ and we assume in the following that all representations are finite-dimensional.

@@

For $m\ge1$ and $1\le i\le n-1$, we write $D_{m,i}$ for the $i$-th copy of $D_m$ inside $D_{m+1}$.
Note that for $1\le r\le n-1$, the maximal Dyck paths $D_{m,i}$, $r+1\le i\le n$, naturally identify with subpaths of $D_{m+1}^{[r]}$.
Extending the notation above, for $m\ge2$ and $1\le r\le n-1$, we write $D_{m,i}^{[r]}$ for the Dyck path obtained by removing the first $r$ copies of $D_{m-1}$ from $D_{m,i}$.
\begin{remark}
  For notational convenience, we also set $D_{m,n}^{[1]}:=D_{m+1}^{[n-1]}$ even though there is no maximal Dyck path $D_{m,n}$ identifying with a copy of $D_m$ inside $D_{m+1}$, such notation is justified by Theorem~\ref{th:dyck path recursion}.
  This should be compared with Corollary~\ref{cor:truncated preprojective isomorphism} and Lemma~\ref{le:truncated quotients}.

  This allows to write $D_{m,n}^{[r]}$ for $1\le r\le n-1$ for the terminal subpaths of $D_{m+1}$.
  We also iterate this notation below by identifying $D_{m,r+1}$ with $D_m$ and identifying $D_{m,r+1,n}^{[r+1]}$ with the subpath obtained by removing the first $r+1$ copies of $D_{m-2}$ from a copy of $D_{m-1}$.
\end{remark}

@@

%\sayT{references to be unified}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{10}

\bibitem{ass} 
  I.~Assem, D.~Simson, A.~Skowronski: Elements of the Representation Theory of Associative Algebras. Cambridge University Press, Cambridge 2007.
\bibitem{ars} 
  M.~Auslander, I.~Reiten, S.~O.~Smalo: Representation theory of Artin algebras {\bf 36}. Cambridge University Press, Cambridge 1997.
\bibitem{bb} 
  A.~Bia\l{}ynicki-Birula: Some theorems on actions of algebraic groups. Annals of Mathematics \textbf{98}, 480-497 (1973).
\bibitem{bgp} 
  I.~N.~Bernstein, I.~M.~Gelfand, V.~A.~Ponomarev: Coxeter functors, and Gabriel's theorem. Russian Mathematical Surveys \textbf{28}(2), 17-32 (1973).
\bibitem{brenner-butler} 
  S.~Benner, M.~C.~R.~Butler: The equivalence of certain functors occuring in the representation theory of artin algebras and species. Journal of the London Mathematical Society \textbf{2}(1), 183-187 (1976).
\bibitem{cc} 
  P.~Caldero, F.~Chapoton: Cluster algebras as {H}all algebras of quiver representations. Commentarii Mathematici Helvetici \textbf{81}(3), 595-616 (2006).
\bibitem{ck} 
  P.~Caldero, B.~Keller: From triangulated categories to cluster algebras II.  Annales Scientifique de l'Ecole Normale Superi\'{e}ure (4) \textbf{39} (6), 983-1009 (2006).
\bibitem{ce} 
  G.~Cerulli Irelli, F.~Esposito: Geometry of quiver Grassmannians of Kronecker type and applications to cluster algebras. Algebra \&  Number Theory \textbf{5}(6), 777-801 (2011).
\bibitem{cefr} 
  G.~Cerulli Irelli, F.~Esposito, H.~Franzen, M.~Reineke: Topology of Quiver Grassmannians. Preprint 2018.
\bibitem{cr} 
  P.~Caldero, M.~Reineke: On the quiver Grassmannian in the acyclic case. Journal of Pure and Applied Algebra \textbf{212}(11), 2369-2380 (2008).
%\bibitem{cj} Crawley-Boevey, W., Jensen, Bernt Tore: A note on sub-bundles of vector bundles. Glasgow Mathematical Journal \textbf{48}, 459-462 (2006).
\bibitem{gab} 
  P.~Gabriel: The universal cover of a finite-dimensional algebra. Representations of algebras. Lecture Notes in Mathematics {\bf 903}, 68-105 (1981).
%\bibitem{hr} 
 % D.~Happel, C.~M.~Ringel: Tilted Algebras. Transactions of the American Mathematical Society {\bf 274}(2), 399-443 (1982).
\bibitem{llz} 
  K.~Lee, L.~Li, A.~Zelevinsky: Greedy elements in rank 2 cluster algebras. Selecta Mathematica \textbf{20}(1), 57-82 (2014).
\bibitem{lw} 
  O.~Lorscheid, T.~Weist: Representation type via Euler characteristics and singularities of quiver Grassmannians. Preprint 2017, arXiv:1706.00860.
\bibitem{rin1} 
  C.~M.~Ringel. Exceptional modules are tree modules. Linear algebra and its Applications \textbf{275/276}, 471-493 (1998).
%\bibitem{rin} 
 % C.~M.~Ringel: Reflection functors for hereditary algebras. Journal of the London Mathematical Society (2) {\bf 21}(3), 465-479 (1980).
%\bibitem{mats} Matsumura, H.: Commutative ring theory. 
\bibitem{rupel} 
  D.~Rupel: Rank Two Non-Commutative Laurent Phenomenon and Pseudo-Positivity. Preprint 2017, arXiv:1707.02696.
\bibitem{sch} 
  A.~Schofield: General representations of quivers. Proceedings of the London Mathematical Society (3) \textbf{65}(1), 46-64 (1992).
\bibitem{wei} 
  T.~Weist: Localization of quiver moduli spaces. Representation Theory \textbf{17}(13), 382-425 (2013).
%\bibitem{wei2} 
 % T.~Weist: Tree modules. Bulletin of the London Mathematical Society \textbf{44}(5), 882-898 (2012).
%\bibitem{wei3} Weist, T.: On the recursive construction of indecomposable quiver representations. Journal of Algebra \textbf{443}, 49-74 (2015).
\end{thebibliography}
